---
title: "Modeling Exercises C"
author: "Suzy An"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part C: Monte Carlo Modeling Exercise

We will be using monte carlo simulations to look at a hypothetical case study of a child's exposure to pesticide from eating fruit.

The risk question that we are interested in answering with our case study  is how much pesticide exposure do children (ages 3–5) receive from ingesting fruit? By entering values for the pesticide concentration in fruit, the child’s daily fruit consumption, and the child’s body weight, we obtain a single value for a child’s ingested dose of pesticide. This approach is limited in that it does not reflect the variability in exposure between different children. 

How much pesticide exposure do children (ages 3–5) receive from ingesting fruit? A simple formula for calculating a child’s ingested dose of pesticide is represented by:

Ingested Dose = (Pesticide Concentration x Fruit Consumption)/ Child Body Weight

In which:

•	Pesticide concentration in fruit is measured in micrograms per gram (μg/g) and is assumed to have a lognormal distribution with a mean of 90 μg/g and a standard deviation of 65 μg/g. 
•	Children’s daily fruit consumption  is measured in grams per day (g/day) and is assumed to have a lognormal distribution with a mean of 236 g/day and a standard deviation of 204 g/day
•	Children’s body weight is measured in kilograms (kg) and is assumed to have a lognormal distribution with a mean of 18 kg and a standard deviation of 4 kg.
•	The ingested dose of pesticide is measured in micrograms per kilogram per day (μg/kg/day).

We can estimate these ingested dose values by making draws from the distributions of pesticide concentration, consumption, and child weight, calculating the resulting Ingested Dose, and repeating many times to get a distribution of ingested dose values. 

## Creating the distributions

Before moving onwards, you're going to have to download the R package pastecs

Let's start by generating the normal distributions of pesticide concentration, consumption, and child weight.

```{r distributions}
set.seed(1)

pesticide <- rnorm(n = 1000, mean = 90, sd = 65)
consumption <- rnorm (n = 1000, mean = 236, sd = 204)
child_weight <- rnorm (n = 1000, mean = 18, sd = 4)

hist(pesticide)
hist(consumption)
hist(child_weight)

summary(pesticide)
summary(consumption)
summary(child_weight)
```

Then let's calculate the resulting Ingested Dose. 

```{r normal}
set.seed(1)

p <- sample(pesticide, 1000)
c <- sample(consumption, 1000)
w <- sample(child_weight, 1000)

ingested_dose <- (p*c/w)

hist(ingested_dose)
summary(ingested_dose)

library(pastecs)
options(scipen = 100)
options(digits = 2)
stat.desc(ingested_dose)
```

Does it make sense for the Ingested Dose to be negative? 
Let's now generate the original lognormal distributions of pesticide concentration, consumption, and child weight.

```{r lognormal}
set.seed(1)

m <- 90
s <- 65
mp<- log(m^2 / sqrt(s^2 + m^2))
sp <- sqrt(log(1 + (s^2 / m^2)))
pesticide_log <- rlnorm(n = 1000, mp, sp)

m <- 236
s <- 204
mc <- log(m^2 / sqrt(s^2 + m^2))
sc <- sqrt(log(1 + (s^2 / m^2)))
consumption_log <- rlnorm(n = 1000, mc, sc)

m <- 18
s <- 4
mw <- log(m^2 / sqrt(s^2 + m^2))
sw <- sqrt(log(1 + (s^2 / m^2)))
child_weight_log <- rlnorm(n = 1000, mw, sw)

hist(pesticide_log)
hist(consumption_log)
hist(child_weight_log)

summary(pesticide_log)
summary(consumption_log)
summary(child_weight_log)
```

Then let's calculate the resulting Ingested Dose. 

```{r ingested_dose}
set.seed(1)

p_log <- sample(pesticide_log, 1000)
c_log <- sample(consumption_log, 1000)
w_log <- sample(child_weight_log, 1000)

ingested_dose_log <- (p_log*c_log/w_log)

hist(ingested_dose_log)
summary(ingested_dose_log)

library(pastecs)
options(scipen = 100)
options(digits = 2)
stat.desc(ingested_dose_log)
```

Let's think about the percentages of children that will have higher exposure values than the crude estimate you calculated in Question 1 (1180 ug/kg/day). 

```{r percentiles}
set.seed(1)

quantile(ingested_dose_log)

quantile(ingested_dose_log, probs = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0))

quantile(ingested_dose_log, probs = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.68, 0.7, 0.8, 0.9, 1.0))

quantile(ingested_dose_log, probs = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.91, 1.0))
```



